---
title: "Iris Data Set"
author: "Eric Leung"
date: "July 28th, 2015"
output: html_document
---

# Introduction

The Iris Data Set was downloaded from the [UCI Machine Learning Repository](http://archive.ics.uci.edu/ml/datasets/Iris).

"This is perhaps the best known database to be found in the pattern recognition
literature. Fisher's paper is a classic in the field and is referenced
frequently to this day. (See Duda & Hart, for example.) The data set contains 3
classes of 50 instances each, where each class refers to a type of iris plant.
One class is linearly separable from the other 2; the latter are NOT linearly
separable from each other."

## Attribute Information

1. sepal length in cm
2. sepal width in cm
3. petal length in cm
4. petal width in cm
5. class:
    - Iris Setosa
    - Iris Versicolour
    - Iris Virginica

# Goal

The iris data set contains four potential predictive values: sepal length, sepal
width, petal length, and petal width. One or more combinations of these
attributes may be used to predict the iris class, which can be one of three
classes: Setosa, Versicolour, or virginica.

Before we can predict on the data, we will perform an exploratory data analysis
to understand nuances in the data.

# Load Packages

```{r}
require(ggplot2)
require(dplyr)
require(GGally)
```

# Load Data

```{r}
dat <- read.csv(file = "../data/iris.data")

# name columns based on attributes above
names(dat) <- c(
    "sepal.length", "sepal.width",
    "petal.length", "petal.width",
    "class")

dat <- tbl_df(dat)
dat
```

# Explore Data

```{r}
str(dat) # take a look at data and kinds of data
```

There appears to be `r dat %>% nrow()` samples with `r dat %>% ncol()` different
variables. Four out of the five appear have numerical data types while the last
variable is a factor variable. This factor variable is what we will predict on,
which is the class of the flower.

```{r}
summary(dat) # look at some summary statistics of data
dat %>% apply(MARGIN = 2, FUN = sd) # look at standard deviation of data
dat %>% apply(MARGIN = 2, FUN = range) # look at range of values vars can take
```

These summary statistics give us a brief overview of how the data looks like.
Each of the different variables do not appear to vary much in range. Petal
length appears to have the greatest standard deviation and widest range.
Additionally, each class of iris are almost equal in frequency. It is also worth
noting there doesn't appear to be any missing values in terms of class.

```{r}
ggpairs(dat) # look at all pairwise plots of variables
```

Here we can look at big picture trends. From here, we might be able to make the
following claims:

- Sepal length appears to be a good distinguisher among the three iris classes
as their box plots do not overlap very much, if any. However, looking at a
different view of the boxplots in terms of histograms, we can see the amount of
overlap among the flowers.
- Petal length and petal width are another variable that might be helpful in
distinguishing among the different classes. Both of their histograms show that
one class of iris, setosa, can be uniquely identified using these two variables.

## Explore Sepal Properties

```{r}
qplot(x = dat$sepal.length,
      main = "Sepal Length Distribution")
qplot(x = dat$sepal.width,
      main = "Sepal Width Distribution")
```

## Explore Petal Properties

```{r}
qplot(x = dat$petal.length,
      main = "Petal Length Distribution")
qplot(x = dat$petal.width,
      main = "Petal Width Distribution")
```
